JAGS Model

### 2020-02-16 Rcodes/JAGS_overdoses.Rmd
JAGS DEBUG DONE!!!!!!!!!!!!!!!!!!!!!!!!!!!!!

---
title: "JAGS_overdoses"
author: "Tom Park"
date: '2020-02-08'
output: html_document
---
# Basic Model: Ambulance + Overdose
## Ambulance Call-outs Model
$n_{A}$: sample size   
$x_{A}$: the total number who confirmed they did call an ambulance  
$p_{A}$: probability of a person call an ambulance   
$$x_{A} \sim Bin(n_{A},p_{A})$$
**We assume $n_{A}=1000, p_A = 0.8$.**  
**Suppose the prior of $p_A$ is noninformative.** 
$$p(p_A) \sim Beta(1,1)$$
## Overdose Model
Now we plug in this values into the overdose model and obtain possible $O_t$ values **assuming we have $U_t$ values.**

@@

```{r message=FALSE, warning=FALSE}
# install packages
install.packages('runjags')
install.packages('rjags')
install.packages('coda')
library('runjags')
library('rjags')
library('coda')
```
The data is the same data from pymc3 with Python.
```{r}
df <- read.csv('./basic_data.csv',check.names=FALSE)
head(df)
```
Now we set the model which defines the relations of overdose model and ambulance call model.
```{r Enter the models}
cat("model{
## define the priors
p_a <- dbeta(alpha, beta)
mu <- dunif(mu_a, mu_b)
sigma <- dunif(sigma_a, sigma_b)

## the latent variables
z <- dnorm(mu, 1/sigma^2)
lmd <-  exp(z)

for (i in 1:N) {
x_a[i] <- dbin(p_a, N) # each survey result for month
o_t[i] <- dpois(lmd) # total overdoses per month
u_t[i] <-dbin(p_a, o_t[i]) # ambulanced overdoses per month
}
}

    ", file='basic_model.txt')
```

```{r Defining Values}
N <-  length(df$o_t) # we have 12 samples
u_t <- df$u_t
x_a <- df$x_a
dat <- list(
            # priors for ambulance model
            'alpha' = 1,
            'beta' = 1,

            # priors for overdose model 
            'mu_a'=-10,
            'mu_b'=0,
            'sigma_a'=0,
            'sigma_b'=5,

            # likelihood 
            'u_t'=u_t, 
            'x_a'=x_a, # We do not include o_t since unknown
            "N"=N # total months
            )

```

I referred two tutorials online. 
[first one](https://www.youtube.com/watch?v=F9jxvhXc_1w) used run.jags for getting posterior distribution and [the second one](https://rstudio-pubs-static.s3.amazonaws.com/272658_ae4d482c86514674be17042c852ebbfc.html) used jags.model function.

I tried both and they didn't work.

This is using jags.model function from rjags package.
```{r Set Up the JAGS Model:jags.model}
jags.simple <- jags.model(file='basic_model.txt',
                          data=dat,
                          n.chains=2,
                          n.adapt=1000)
```

# Q1.

Here I am confused since dunif is what I found correct [here](https://web.sgh.waw.pl/~atoroj/ekonometria_bayesowska/jags_user_manual.pdf).

# Q2. 
When I tried the other package and function, I got an error that object "NA" not found as below.

This is using jags.model function from runjags package.
```{r another try:run.jags}
jags.simple <- run.jags(file='basic_model.txt',
                          data=dat,
                          monitor =c('u_t'),
                          n.chains=2,
                          n.adapt=1000)
```
```

>>>>

---
title: "JAGS Model"
author: "Tom Park"
date: '2020-02-08'
output: html_document
---
## Basic Model: Ambulance + Overdose
### Ambulance Call-outs Model
$n_{A}$: sample size   
$x_{A}$: the total number who confirmed they did call an ambulance  
$p_{A}$: probability of a person call an ambulance   
$$x_{A} \sim Bin(n_{A},p_{A})$$
**We assume $n_{A}=1000, p_A = 0.8$.**  
**Suppose the prior of $p_A$ is noninformative.** 
$$p(p_A) \sim Beta(1,1)$$
### Overdose Model

Now we plug in this values into the overdose model and obtain possible $O_t$ values **assuming we have $U_t$ values.**

@@

```{r message=FALSE, warning=FALSE}
# install packages
if (!require(runjags)) install.packages("runjags")
if (!require(rjags)) install.packages("rjags")
if (!require(coda)) install.packages("coda")

library('runjags')
library('rjags')
library('coda')
```
The data is the same data from pymc3 with Python.  
Todo: build a pipeline to connect the python (pymc3) and R (JAGS)

```{r}
df <- read.csv('./basic_data.csv')
df$X <- NULL 
head(df)
```
Now we set the model which defines the relations of overdose model and ambulance call model.

The model defined as follows.
```{r Enter the models}
cat("model{
## define the priors
p_a ~ dbeta(alpha, beta)
mu ~ dunif(mu_a, mu_b)
sigma ~ dunif(sigma_a, sigma_b)

## the latent variables
z ~ dnorm(mu, 1/(sigma^2))
lambda <- exp(z)

for (i in 1:n) {
  ## ambulance model 
  x_a[i] ~ dbin(p_a, n_a) # each survey result for month
}
for (i in 1:n) {
  ## overdose model
  o_t[i] ~ dpois(lambda*N) # total overdoses per month
}
for (i in 1:n) {
  u_t[i] ~ dbin(p_a, o_t[i]) # ambulanced overdoses per month
}

}", file='basic_model.txt')
```

Pre-set variables.
```{r Defining Values}
n <-  length(df$o_t) # we have 12 samples
n_a <- 1000
N <- 10000
u_t <- df$u_t
x_a <- df$x_a
```

Define the list providing the values of the variables and the parameters for the priors of the model. 
```{r from variable name to \'variable_name\'}
dat <- list(
            # priors for ambulance model
            'alpha' = 1,
            'beta' =  1,

            # priors for overdose model 
            'mu_a'=(-10),
            'mu_b'=0,
            'sigma_a'=0,
            'sigma_b'=5,

            # likelihood 
            'u_t'=u_t, 
            'x_a'=x_a,  
            'N'=N, # the population
            'n'= n, # total months
            'n_a'=n_a

            )
```
**Note: for the list object usually named 'data' or 'dat' in JAGS context, do not use arrow but use equal sign to define elements of the list.**
``` {r run the first model without error }
jags.simple <- jags.model(file='basic_model.txt',
                          data=dat)
```

## Reference 
[first tutorial](https://www.youtube.com/watch?v=F9jxvhXc_1w)  
[second tutorial](https://rstudio-pubs-static.s3.amazonaws.com/272658_ae4d482c86514674be17042c852ebbfc.html)   
[JAGS manual](https://web.sgh.waw.pl/~atoroj/ekonometria_bayesowska/jags_user_manual.pdf)   
[error handling guide](https://www4.stat.ncsu.edu/~reich/ABA/code/errors)
